---
title: "Media Franchises"
author: "Randall Madsen"
date: "July 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(tidyverse)
library(scales)
library(plotly)

media_franchises <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-02/media_franchises.csv")

theme_set(theme_classic())
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
graph_1 <- ggplotly(media_franchises %>%
  filter(str_detect(owners, "Disney")) %>% 
  group_by(franchise) %>% 
  mutate(total_revenue = sum(revenue)) %>% 
  distinct(franchise, .keep_all = T) %>% 
  ungroup() %>% 
  mutate(franchise = fct_reorder(franchise, total_revenue)) %>% 
  ggplot(aes(franchise, total_revenue, fill = original_media,
             text = paste('Franchise:', franchise,
                          '</br>',
                          '</br> Total Revenue Amount: $',total_revenue))) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(x = "",
       y = "Total Revenue (in Billions)",
       fill = "",
       title = "Media Franchises Owned by The Walt Disney Company"),
  tooltip = "text")

```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}



graph_2 <- ggplotly(media_franchises %>%
  filter(str_detect(owners, "Disney")) %>% 
  group_by(franchise) %>% 
  mutate(total_revenue = sum(revenue)) %>% 
  ungroup() %>% 
  mutate(franchise = fct_reorder(franchise, total_revenue)) %>% 
  ggplot(aes(franchise, revenue, fill = revenue_category,
             text = paste('Franchise:', franchise,
                          '</br>',
                          '</br> Revenue Amount: $',round(revenue),
                          '</br> Revenue Category: ', revenue_category))) +
  geom_col(position = "stack") +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme(legend.position = "none") +
  coord_flip() +
  # facet_wrap(~franchise, scales = "free") +
  labs(fill = "", 
       x = "",
       y = "",
       title = "Media Franchises Owned by The Walt Disney Company",
       subtitle = "Broken down by Revenue Category"),
  tooltip = "text") 
 

# axis.title.x=element_blank(),
#         axis.text.x=element_blank(),
#         axis.ticks.x=element_blank(),


  # layout(title = "New plot title",
  #    #xaxis = list(showticklabels = FALSE),
  #    legend = list(orientation = "v"))
```
 
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
distinct_top_revenue_source <- media_franchises %>% 
  group_by(franchise) %>% 
  mutate(top_revenue = max(revenue)) %>% 
  ungroup() %>% 
  filter(top_revenue == revenue)

graph_3 <- ggplotly(distinct_top_revenue_source %>%  
  ggplot(aes(as.factor(revenue_category),  revenue, fill = franchise,
             text = paste('Franchise:', franchise,
                          '</br>',
                          '</br> Revenue Amount: $',round(revenue)))) +
  geom_point(aes(col = franchise), size=3) +
  theme(legend.position="none") +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(x = "",
       y = "Revenue (in Billions)",
       title = "Top Revenue Source by Franchise") +
  coord_flip(),
  tooltip = "text")

```

##This was another attempt at a [tidytuesday](https://github.com/rfordatascience/tidytuesday) blog post. This [dataset](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-07-02) focused on media franchises. The first one that I was curious about was how The Walt Disney Company stacked up by total revenue. 
    
### The Walt Disney Company (Total Revenue)
    
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
graph_1
```

## I then wanted to see how their revenue was divided between the different sources. 
    
### The Walt Disney Company (Revenue by Source)
    
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
graph_2
```

## After exploring Disney franchises, I was curious about the overall top revenue source by franchise.

### Top Revenue Source by Franchise 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
graph_3
```

## It is interesting to see that there are `r nrow(distinct_top_revenue_source %>% filter(revenue_category == "Merchandise, Licensing & Retail"))` total in the Merchandise, Licensing & Retail category. This equates to `r round((nrow(distinct_top_revenue_source %>% filter(revenue_category == "Merchandise, Licensing & Retail")))/ (nrow(distinct_top_revenue_source)) * 100, 1)` % of the top total highest revenue source. 

## In summary, generating revenue through merchandise, licensing and retail will generally provide the most for franchises.  